### 2019-03-08 thesis/abstract.tex
moving thesis directory

%
% Abstract
%

\thispagestyle{empty}
\begin{center}

{\Large \bf ABSTRACT}

\vskip.35in
{\Large \bf TITLE HERE}

\vskip.35in
{\large YOUR NAME} \\
\vskip.35in
\end{center}
In this \thesis... TEXT HERE