### 2019-03-08 thesis/abstract.tex
title and name

%
% Abstract
%

\thispagestyle{empty}
\begin{center}

{\Large \bf ABSTRACT}

\vskip.35in
{\Large \bf The Feedback-Driven Evolution of Metals in Dwarf Galaxies as Simulated with Individual Stars}

\vskip.35in
{\large Andrew J. Emerick} \\
\vskip.35in
\end{center}
In this \thesis...