### 2019-03-19 thesis/acknowledgments.tex
adding year and location

%
%
% Acknoledgements
%
%

\newpage

\begin{center}

{\large \bf ACKNOWLEDGMENTS } % leave this title

\end{center}

\vspace{0.8cm}
% Thank people, places, things, existential concepts, etc. here. Length varies
% but usually something like 1-2 pages
\textit{To-do}

\vspace{1.8cm}
2019, New York, NY

%\setlength{\baselineskip}{1.11111 \baselineskip}