### 2019-04-08 thesis/acknowledgments.tex
adding software

%
%
% Acknoledgements
%
%

\newpage

\begin{center}

{\large \bf ACKNOWLEDGMENTS } % leave this title

\end{center}

\vspace{0.8cm}
% Thank people, places, things, existential concepts, etc. here. Length varies
% but usually something like 1-2 pages
\textit{To-do}

To my wife Ellen, who moved across the country for me to N.Y.C. and has put up
with so much over the past six years just so that I could pursue my dream.

I want to thank all of the other graduate students at Columbia that I have overlapped
with over the years. This includes especially Sarah, Steven, Dan, Jeff, Adrian,
Munier, Andrea, Alejandro, and Susan. \textit{a little more here}

Thank you to everyone I have met in the NYC climbing community who have collectively
help keep me sane throughout my time in graduate school. Over the past few years
I have learned the importance of a proper work-life balance, and climbing has helped
me keep that balance (... maybe a bit too well...). Thank you especially to my
regular climbing partners -- Daniel, Hunter, Dennis, and Lauren A. -- who I've trusted
with my life countless times, and the handful of other astronomer-climbers that I've
climbed with throughout graduate school (Jeff, Matt, Lauren A., Dan D'Orazio,
Dan Foreman-Mackey, and Andrea).

Finally, I would like to thank the numerous developers and contributors to the
open-source software packages without which this \dissertation would not be
possible. This includes more general software, such as \textsc{Python}, \textsc{IPython}, \textsc{NumPy},
\textsc{SciPy}, \textsc{Matplotlib}, \textsc{HDF5}, \textsc{h5py},
and \textsc{deepdish}, and more astronomy-specific projects like  \textsc{yt}, \textsc{Enzo},
\textsc{Grackle}, \textsc{AstroPy}, and \textsc{Cloudy}. In particular I would like
to thank the many developers of \textsc{yt}, \textsc{Enzo}, and \textsc{Grackle}
who have helped answer questions, debug code, and implement new features on
my behalf throughout this process. This includes: Britton Smith, Nathan Goldbaum,
John Wise, Matt Turk, Greg Bryan, Simon Glover, Munier Salem, Cameron Hummels,
Christine Simpson, and Brian O'Shea.

\vspace{1.8cm}
2019, New York, NY

%\setlength{\baselineskip}{1.11111 \baselineskip}